---
output: rmarkdown::github_document
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse=TRUE, 
                      comment="##", 
                      fig.retina=2, 
                      fig.path = "README_figs/README-")
```

## ggerp: Graphical exploration of ERP data with R

<!-- README.md is generated from README.Rmd. Please edit that file -->

**ggerp** considers event-related brain potential (ERP) data analysis in a standard linear model setting and implements a set of functions in R to facilitate visual exploration of data and display of statistical testing results.

These R functions are based on [ggplot2 package](https://cran.r-project.org/web/packages/ggplot2/ggplot2.pdf) to enable step-by-step revision of graphic objects. We also adopts significant testing procedure provided by [ERP package](https://cran.r-project.org/web/packages/ERP/index.html).

Beyond rendering conditioning plots of ERPs on the scalp map, with these functions in `ggerp` users can create animation for displaying significant effects across electrode locations over time by using [gganimate package](https://github.com/dgrtwo/gganimate). 

The `ggerp` can be installed using [devtools](https://github.com/hadley/devtools):

```
install_github("PsyChiLin/ggerp")
```

**Packages Preparation**

The following R command lines illustrate the exploration of ERP data using advanced graphical tool `ggerp` available in R.

Load (and install) packages. 
```{r pkgs,message=FALSE}
library(pacman)
pacman::p_load(ERP, mnormt, fdrtool, tidyverse, gridExtra, crayon,
               boot, reshape2, ggthemes, devtools)
```

Install `ImageMagick` within R. Please check all boxes.
```{r pkgs_IM,message=FALSE}
#install.ImageMagick()
```

Install and load `gganimate` and `animation` from github.
```{r pkgs_GGan,message=FALSE}
#install_github("yihui/animation")
#install_github("dgrtwo/gganimate")
library(animation)
library(gganimate)
```

Load `ggerp`.
```{r pkgs_ggerp,message=FALSE}
library(ggerp)
```
**Data Preparation**
We demonstrate the graphical capabilities of `ggerp` with these real data set `DirectedForgetting`. Download data `DirectedForgetting` from [this website](https://www.dropbox.com/s/20uhxsmcbex3i0m/DirectedForgetting.csv?dl=0) or directly use the build-in one.
It contains variables named `TBR_score` and `TBF_score`  (continuous), `Condition` (categorical), and one variable per time point (ERP values, i.e, `T_1200`). The scope of possible linear modeling designs is therefore quite large. The command lines shall be marginally adpated to your own ERP dataset. Note that `Condition`, which is a with-subject variable, could also be changed to a between subject variable `Group` in your own dataset.
```{r data,message=FALSE}
# dta <- read.csv("DirectedForgetting.csv")
dta <- DirectedForgetting
```

The sequence of time points is generated, called `time_pt`.
```{r pt,message=FALSE}
time_pt <- seq(-200, 1000, 1)
```

Specify channels in the data file `erpR_coord` according to their scalp locations.
```{r coord,message=FALSE}
erpR_coord <- rbind(c(NA, "FP1", NA, "FP2", NA),
                    c("F7", "F3", "FZ", "F4", "F8"),
                    c("FT7", "FC3", "FCZ", "FC4", "FT8"),
                    c("T7", "C3", "CZ", "C4", "T8"),
                    c("TP7", "CP3", "CPZ", "CP4", "TP8"),
                    c("P7", "P3", "PZ", "P4", "P8"),
                    c(NA, "O1", "OZ", "O2", NA))
```

**Average ERP curves for conditions on the 10/10 system**
```{r F01p,message=FALSE}
Fig01 <- plot_tete(data = dta,
                   frames = time_pt,
                   channel = 5,
                   subject = 1,
                   uV = 6:1206,
                   test = 4,
                   mode = "mean",
                   scalp = TRUE,
                   curve.col = c("chartreuse4","firebrick"),
                   coord.mat = erpR_coord,
                   ylim = c(-5, 10))
```

**Individual ERP curves by condition from four channels.** 

First, restrict the data to four channels
```{r F02d,message=FALSE}
dta_c <- filter(dta,Channel %in% c("FZ", "FCZ", "CZ", "PZ")) %>%
        droplevels()
```
Then, create the plot.
```{r F02p,message=FALSE}
Fig02 <- plot_tete(data = dta_c,
                   frames = time_pt,
                   channel = 5,
                   subject = 1,
                   uV = 6:1206,
                   test = 4,
                   mode = "raw",
                   ylim = c(-25, 25))
```

**Confidence intervals for mean ERP curves on CZ.** 

First restrict the data to channel CZ.
```{r F03d,message=FALSE}
dta_c <- filter(dta, Channel == "CZ" ) %>%
        droplevels()
```


Seed random number generator for replication
```{r F03s,message=FALSE}
set.seed(123)
```

Then, create the plot in two stages.
```{r F03p,message=FALSE}
Fig03 <- plot_tete(data = dta_c,
                   frames = time_pt,
                   channel = 5,
                   subject = 1,
                   uV = 6:1206,
                   test = 4,
                   mode = "bootci",
                   ylim = c(-7, 10))+
        theme(legend.position = c(.9, .9))
```


**Results of significant testing comparing conditions for three channels**

First restrict the data to three channels.
```{r F04d,message=FALSE}
dta_c <- dta %>%
        filter(Channel %in% c("FZ","CZ","PZ")) %>%
        droplevels()
```
Then, create the plot.
```{r F04,message=FALSE, fig.height=3}
Fig04 <- plot_fa(data = dta_c,
                 frames = time_pt,
                 channel = 5,
                 subject = 1,
                 uV = 6:1206,
                 test = 4,
                 mode = "test_signal",
                 design = (~Subject + Condition),
                 design0 = (~Subject),
                 nbf = 5,
                 ylim = c(-6, 7.5))
```

**Significant changes across channels over time on the scalp**

First, save test results.
```{r F05d,message=FALSE, eval=FALSE}
test_res <- plot_fa(data = dta,
                    frames = time_pt,
                    channel = 5,
                    subject = 1,
                    uV = 6:1206,
                    test = 4,
                    mode = "test_signal",
                    design = (~Subject + Condition),
                    design0 = (~Subject),
                    nbf = 5,
                    ylim = c(-6, 13))
```

Then, create GIF file for animation. The example Gif file could be download on [this website](https://www.dropbox.com/s/p80k9jmbburzynv/Fig05.gif?dl=0).
```{r F05p,message=FALSE,eval=FALSE}
Fig05 <- plot_coord(tests_rst = test_res$Test_Rst,
                    frames = time_pt,
                    show = seq(200, 500, by = 1),
                    loop = 1,
                    interval = 0.1, 
                    filename = "Fig05.gif")
```

**Associations between ERPs and a numerical covariate**

First restrict the data to three channels.
```{r F06d, message=FALSE, fig.show = 'hide'}
dta_c <- dta %>%
        filter(Condition == "TBF", Channel %in% c("FZ","CZ","PZ")) %>%
        droplevels()
```

Produce each plot respectively. 
```{r F06p, message=FALSE, fig.show = 'hide'}
Fig06a <- plot_tete(data = dta_c,
                    frames = time_pt,
                    channel = 5,
                    subject = 1,
                    uV = 6:1206,
                    test = 3,
                    mode = "raw",
                    ylim = c(-25, 28))
Fig06a <- Fig06a + 
                theme(legend.position = "top", legend.box = "horizontal")
Fig06b <- plot_tete(data = dta_c,
                    frames = time_pt,
                    channel = 5,
                    subject = 1,
                    uV = 6:1206,
                    test = 3,
                    mode = "bootci",
                    ylim = c(-1, 1),
                    labs = list(y = "Correlation", x = "Time (ms)"))
Fig06c <- plot_fa(data = dta_c,
                  frames = time_pt,
                  channel = 5,
                  subject = 1,
                  uV = 6:1206,
                  test = 3,
                  mode = "test",
                  design = (~TBF_score),
                  nbf = 5,
                  ylim=c(-1, 1),
                  labs = list(y = "Correlation", x = "Time (ms)"))
```
Combine the plots to a final figure.
```{r F06all,message=FALSE, fig.height=10}
grid.arrange(Fig06a +
                     ggtitle("A") +
                     theme(plot.title = element_text(hjust = 0)),
             Fig06b +
                     ggtitle("B") +
                     theme(plot.title = element_text(hjust = 0)),
             Fig06c$Plot +
                     ggtitle("C") +
                     theme(plot.title = element_text(hjust = 0)))
```
